---
title: "stan examples"
author: "Yanruyu Zhu (yaz4004)"
date: "12/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits=6)
```

This file contains examples of 'Stan'.         
The first example is for Gaussian model.      
The second example is for Beta-binomial model.

 
```{r message=FALSE, include=FALSE}
#remove.packages("rstan")
#if (file.exists(".RData")) file.remove(".RData")
#install.packages('rstan')
```

## Set up

```{r message=FALSE}
library('rstan')
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
seed = 2021
```

## Ex_1 Gaussian    

The mock data is 1,000 i.i.d. X ~ Normal($\mu$=5,$\sigma$=1).      

```{r message=FALSE}
X = rnorm(n=1000, mean=5, sd=1) # fake data
mydata = list(N=1000, X=X)      # stan takes data format as 'List'
fit = stan(file='my_model_norm.stan', data=mydata, seed = seed)
#print(fit)
```

### Template Result:     

```{r}
print(fit)
```
### Pair plots:      

```{r fig.cap="Pair plots for Gaussian model"}
pairs(fit)
```

### Try to extract sample results

```{r}
param = extract(fit)
param$mu[1:10]
```

```{r, include=FALSE}
# could re-load after clear workspace 
save(fit, file='my_fit')
```

```{r, include=FALSE}
# reload previously saved my_fit
load('my_fit')
print(fit)
```

## Ex_2 Beta-binomial        

### Parameters and set up:       

The number of sampling times is 1000;      
Prior distribution $\theta$~$Beta(1,1)$, Y~$Bin(100,\theta)$;         
Observed data is 82 responses in the first cohort of 100 patients;        

```{r message=FALSE}
iter = 1000
data_bin_low <- list(N = 100, y = 82)

fit_bin_low <- stan(file = 'beta_binom.stan', data = data_bin_low, iter = iter, seed = seed)
```

### Posterior distribution for $\theta$:       

```{r}
monitor(fit_bin_low)
draws <- as.data.frame(fit_bin_low)
hist(draws[,'theta'])
```

### For $P(\theta \leq 0.92 | Y=82)$ 

```{r}
sum(draws$theta<=0.92)/nrow(draws)
```

As $P(\theta \leq 0.92 | Y=82) = 0.9985 >0.9$, we stop the trial at the first interim analysis for observing 82 responses in the first 100 patients.


```{r}
#iter = 1000
data_bin_h <- list(N = 100, y = 95)

fit_bin_h <- stan(file = 'beta_binom.stan', data = data_bin_h, iter = iter, seed = seed)
```

```{r}
draws_h <- as.data.frame(fit_bin_h)
sum(draws_h$theta<=0.92)/nrow(draws_h)
```

### Iteration from Y = 82 to Y = 95       

```{r}

p_92 = rep(0, 95-81)

for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

```{r}
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```


```{r}
paste("P|Y=88:", p[88-81])
paste("P|Y=89:", p[89-81])
```

### j

```{r}
a = 1
b = 1
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

mean = 0.9, roughly same variance as Beta(1,1)

```{r}
a = 0.090
b = 0.010
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

```{r}

p = rep(0, 95-81)
iter = 1000
for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

```{r}
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```

```{r}
paste("P|Y=88:", p[88-81])
paste("P|Y=87:", p[87-81])
```

