---
title: "Prior distribution study"
author: "Yanruyu Zhu (yaz4004)"
date: "1/09/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits=6)
```

```{r message=FALSE, include=FALSE}
#remove.packages("rstan")
#if (file.exists(".RData")) file.remove(".RData")
#install.packages('rstan')
```

## Set up               
Seed = 2021;             
The number of sampling times is 1000;

```{r message=FALSE}
library('rstan')
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
seed = 2021
iter = 1000
```

## Parameters      
      
Prior distribution $\theta$~$Beta(\alpha,\beta)$, Y~$Bin(100,\theta)$;         
    

### Scenario 1:                 
$\alpha = 1, \beta = 1$            

- Mean and Variance of Beta(1,1)           

```{r}
a = 1
b = 1
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

```{r message=FALSE}
p = rep(0, 95-81)

for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom_1_1.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

- Posterior plot for number of success from 82 to 95          

```{r}
par(mfrow = c(2, 2))
## Plot 1
x_beta <- seq(0, 1, by = 0.025 )  
y_beta <- dbeta(x_beta, shape1 = 1, shape2 = 1)  
# Plot beta values
plot(x_beta, y_beta, type="o", pch="o", lty=1, col="black",
     xlab = "X",
     ylab = "Probability", 
     main = "Pdf of Beta(1,1)")
##################################
## Plot 2
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta",
     sub = "a=1, b=1")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```

- Threshhold value         

```{r}
paste("P|Y=88:", p[88-81])
paste("P|Y=89:", p[89-81])
```

### Scenario 2:                   

$\alpha = 0.9, \beta = 0.1$            

- Mean and Variance of Beta(0.9,0.1)

```{r}
a = 0.9
b = 0.1
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

Mean is 0.9, and variance is relatively high.

```{r message=FALSE}
p = rep(0, 95-81)

for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom_0.9_0.1.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

- Posterior plot for number of success from 82 to 95            

```{r}
par(mfrow = c(2, 2))
## Plot 1
x_beta <- seq(0, 1, by = 0.025 )  
y_beta <- dbeta(x_beta, shape1 = 0.9, shape2 = 0.1)  
# Plot beta values
plot(x_beta, y_beta, type="o", pch="o", lty=1, col="black",
     xlab = "X",
     ylab = "Probability", 
     main = "Pdf of Beta(0.9,0.1)")
##################################
## Plot 2
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta",
     sub = "a=0.9, b=0.1")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```

- Threshhold value         

```{r}
paste("P|Y=87:", p[87-81])
paste("P|Y=88:", p[88-81])
```

### Scenario 3:                   
$\alpha = 9, \beta = 1$            

- Mean and Variance of Beta(9,1)        
```{r}
a = 9
b = 1
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

Mean is 0.9, and variance is relatively medium.

```{r message=FALSE}
p = rep(0, 95-81)

for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom_9_1.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

- Posterior plot for number of success from 82 to 95          

```{r}
par(mfrow = c(2, 2))
## Plot 1
x_beta <- seq(0, 1, by = 0.025 )  
y_beta <- dbeta(x_beta, shape1 = 9, shape2 = 1)  
# Plot beta values
plot(x_beta, y_beta, type="o", pch="o", lty=1, col="black",
     xlab = "X",
     ylab = "Probability", 
     main = "Pdf of Beta(9,1)")
##################################
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta",
     sub = "a=9, b=1")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```

- Threshhold value         

```{r}
paste("P|Y=87:", p[87-81])
paste("P|Y=88:", p[88-81])
```


### Scenario 4:                   
$\alpha = 90, \beta = 10$            

- Mean and Variance of Beta(90,10)          
```{r}
a = 90
b = 10
mean = a/(a+b)
var = (a*b)/((a+b)^2*(a+b+1))
paste("mean = ",mean)
paste("variance = ", var)
```

Mean is 0.9, and variance is way lower.

```{r message=FALSE}
p = rep(0, 95-81)

for (i in 82:95) {
  #a = i+1
  data_bin = list(N = 100, y = i)
  fit_bin = stan(file = 'beta_binom_90_10.stan', data = data_bin, iter = iter, seed = seed)
  draws = as.data.frame(fit_bin)
  p[i-81] = sum(draws$theta<=0.92)/nrow(draws)
}
```

- Posterior plot for number of success from 82 to 95          

```{r}
par(mfrow = c(2, 2))
## Plot 1
x_beta <- seq(0, 1, by = 0.025 )  
y_beta <- dbeta(x_beta, shape1 = 90, shape2 = 10)  
# Plot beta values
plot(x_beta, y_beta, type="o", pch="o", lty=1, col="black",
     xlab = "X",
     ylab = "Probability", 
     main = "Pdf of Beta(90,10)")
##################################
plot(82:95, p, type="o", pch="o", lty=1, col="black",
     xlab = "Number of Success (Y)",
     ylab = "Probability", 
     main = "Posterior distribution of theta",
     sub = "a=90, b=10")
axis(side = 1, at = seq(82, 95, by=1))#, las=2)
abline(h = 0.92, col="red")
legend(x = "topright", legend=c("0.92"),
       col=c("red"), lty=1, lwd=2, cex=0.75)
```

- Threshhold value         

```{r}
paste("P|Y=88:", p[88-81])
paste("P|Y=89:", p[89-81])
```





